---
title: "BaseR"
format: html
editor: visual
---
# Homework 2

## Task 1: Vectors

1. Assign BP values for pre- and post-treatment to vectors
```{r}
bpPre <- c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 127, 135, 133, 128, 147, 138, 140, 132)
bpPost <- c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 120, 114, 121, 120, 128, 120, 133, 115)
```

2. Assign names to the vector elements
```{r}
bpNames <- paste("Subject", 1:20, sep= "_")
names(bpPre) <- bpNames
names(bpPost) <- bpNames
```

3. Assign BP change to new vector (value > 0 indicates decrease)
```{r}
bpChange <- bpPre - bpPost
```

4. Average decrease for all patients
```{r}
mean(bpChange)
```

5. Patients with a decrease in BP
```{r}
which(bpChange > 0)
```

6. Subset of BP changes that are greater than 0
```{r}
bpChange[bpChange > 0]
```

7. Average decrease in BP for patients with positive change
```{r}
mean(bpChange[bpChange > 0])
```

## Task 2: Data Frames

1. create data frame from above data
```{r}
bp_df = data.frame(patient = 1:20, pre_bp = bpPre, post_bp = bpPost,
                   diff_bp = bpChange)
```

2. Rows where diff_bp < 0
```{r}
bp_df[bp_df$diff_bp < 0,]
```

3. Add new column denoting whether post BP is less than 120 or not
```{r}
bp_df$less120 <- bp_df$post_bp < 120
```

4. Print data frame
```{r}
knitr::kable(bp_df)
```

## Task 3: Lists

1. Create new data frame with placebo patients
```{r}
# vectors for data columns
placeboPre <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
placeboPost <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
placeboDiff <- placeboPre - placeboPost 
placebo120 <- placeboPost < 120

# assign names (using subset of bpNames from above)
names(placeboPre) <- bpNames[1:10]
names(placeboPost) <- bpNames[1:10]

# create data frame
placebo_df <- data.frame(patient = 1:10, pre_bp = placeboPre, 
                         post_bp = placeboPost, diff_bp = placeboDiff,
                         less120 = placebo120)

```

2. Create a list for treatment and placebo data
```{r}
bp_list <- list(treatment = bp_df, placebo = placebo_df)
```

3. Access the first list element
```{r}
# syntax 1
bp_list[1]

# syntax 2
bp_list["treatment"]

# syntax 3
bp_list$treatment

```
4. Access just the pre_bp column of the placebo data frame
```{r}
bp_list[["placebo"]]$pre_bp
```



